---
title: "Redlining geospatial notebook"
output: html_notebook
---

```{r}
library(sf)
library(tidyverse)
```

```{r}
HRI2000 <- st_read("data/HRI2000/HRI2000.shp")
```


```{r}
HRI2000 |> 
    ggplot()+
    geom_sf()
```

DOH!!! it's the shape file for the whole effin country!!

```{r}
HRI2000 |> 
    head()
```
filter for davidson county:
```{r}
HRI2000 <- HRI2000 |> 
    filter(str_detect(METRO_NAME, "Davidson"))

```
set HRI levels as a vector:
```{r}
HRI_levels_vec <-c("1"="green", "2"="blue", "3"="yellow", "4"="red")
```
use HRI vector to manually set fill colors for HRI districts  
```{r}
HRI2000 |> 
    ggplot()+
    geom_sf(aes(color = factor(EQINT2000)))+
    scale_color_manual(values = HRI_levels_vec)+
    theme_bw()
```

Now repeat for 2010:

```{r}
HRI2010 <- st_read("data/HRI2010/HRI2010.shp") |> 
    filter(str_detect(METRO_NAME, "Davidson"))

```
```{r}

HRI2010 |> 
    ggplot()+
    geom_sf(aes(fill = factor(EQINTERV10)))+
    scale_fill_manual(values = HRI_levels_vec)
```

Now repeat for 2020:

```{r}
HRI2020 <- st_read("data/HRI2020/HRI2020.shp") |> 
    filter(str_detect(METRO_NAME, "Davidson"))

```
```{r}

HRI2020 |> 
    ggplot()+
    geom_sf(aes(fill = factor(EQINTER20)))+
    scale_fill_manual(values = HRI_levels_vec)
```

Not exactly sure why the delineations are different for these since they are based on historicaly redlining districts and overlayed on 2000, 2010, and 2020 maps.


import University of Richmond geojson:

```{r}
UR_inequality <- st_read('data/mappinginequality.json') |> 
    filter(str_detect(city, "Nashville"))

```
create new vec based on HOLC grade for UR dataset
```{r}
HOLC_grade_vec <-c("A"="green", "B"="blue", "C"="yellow", "D"="red")

```

```{r}
UR_inequality |> 
    ggplot()+
    geom_sf(aes(fill = factor(grade)))+
    scale_fill_manual(values = HOLC_grade_vec)

```

