---
title: "redlining_data_collection"
output: html_notebook
---

Import libraries:

```{r}
library(tidyverse)
library(readxl)
library(sjmisc)
```

Import Redlining indicator data files (excel):
2000
```{r}
Redlining_Indicator_2000 <- read_excel("data/Redlining_Indicator_2000.xlsx")
```

2010
```{r}
Redlining_Indicator_2010 <- read_excel("data/Redlining_Indicator_2010.xlsx")
```

2020
```{r}
Redlining_Indicator_2020 <- read_excel("data/Redlining_Indicator_2020.xlsx")
```

Import census tract income data files (csv):
2000
```{r}
Raw_income_census_2000 <- read_csv("data/DECENNIALDPSF32000.csv")
```

2010
```{r}
Raw_income_census_2010 <- read_csv("data/ACSST5Y2010.S1901-Data.csv")
```

2020
```{r}
Raw_income_census_2020 <- read_csv("data/ACSST5Y2020.S1901-Data.csv")

```

import metadata files for column headers and concat into vectors:
2010:
```{r}
census_headers_2010 <- read_csv("data/ACSST5Y2010.S1901-Column-Metadata.csv")

```
Determine which income measurement and its division levels are present in all three census income data files. According to the census website, Household income takes into account all earnings from each household member but aso income from SS, dividends, interest, and other sources.
Therefore, need to pull household income from each data set so that we are comparing apples to apples.


filter 2010 header file for "household" and remove all other unnecessary rows that aren't household income levels
```{r}
census_headers_2010<-census_headers_2010 |> 
    filter(str_detect(Label, "Household")) |> 
    filter(str_detect(Label, "IMPUTED", negate=TRUE)) |> 
    filter(str_detect(Label, "Margin", negate=TRUE))

```


2020:
```{r}
census_headers_2020 <- read_csv("data/ACSST5Y2020.S1901-Column-Metadata.csv")

```
filter 2020 header file for "household" and remove all other unnecessary rows that aren't household income levels
```{r}
census_headers_2020<-census_headers_2020 |> 
    filter(str_detect(Label, "Household")) |> 
    filter(str_detect(Label, "IMPUTED", negate=TRUE)) |> 
    filter(str_detect(Label, "Margin", negate=TRUE)) |> 
    filter(str_detect(Label, "ALLOCATED", negate=TRUE))
```

compare header files:
```{r}
census_headers_2010 |> 
    anti_join(census_headers_2020, by="Column Name")
```
No stray entries so headers match up

Use resultant headers to create vector to select those columns from raw income datasets.
```{r}

headers_vector_2010<-census_headers_2010 |>
    pull(`Column Name`)
```
add those column names to geo_id and name columns then select columns.
```{r}
headers_vector_2010<- c("GEO_ID", "NAME", headers_vector_2010)
```
```{r}
davidson_income_census_2010<-Raw_income_census_2010 |> 
    select(all_of(headers_vector_2010))
```

2020:
```{r}
headers_vector_2020<-census_headers_2020 |>
    pull(`Column Name`)
```
```{r}
headers_vector_2020<- c("GEO_ID", "NAME", headers_vector_2020)
```
```{r}
davidson_income_census_2020<-Raw_income_census_2020 |>
    select(all_of(headers_vector_2020))
```



Raw income data for 2000 need rows 53-66 which keeps household income for all of the nashville census tracts then to be rotated to long version:
```{r}
davidson_income_census_2000<-Raw_income_census_2000 |> 
    slice(53:66) |> 
    rotate_df(cn=TRUE)

```

census income for 2010 and 2020 have no absolute numbers, just tract population and percentages. need to mutate to show numbers.
```{r}
davidson_income_census_2010 
    #mutate()
```
```{r}
davidson_income_census_2000[1,1]
```
shows that the first column data is an index and not showing up as variables. need to convert the rownames to a column
```{r}
davidson_income_census_2000 <- davidson_income_census_2000 |> 
    rownames_to_column(var="NAME")
```
also this is a df and not a tibble so that needs to be converted:
```{r}
tibble(davidson_income_census_2000)
```
In order to separate out percentages from number rows, we must first split the first column into 2 columns using '!!' as the separator.
```{r}
davidson_income_census_2000 <- davidson_income_census_2000 |> 
    separate_wider_delim(NAME,delim = '!!', names=c("NAME","VALUE"))
```
find column names first
```{r}
davidson_income_census_2000 |> 
    colnames()
```
clean up and rename columns
```{r}
income_headers <- c("NAME", "VALUE", "HOUSEHOLDS", "income<$10k", "income$10-$14,999","income$15-$24,999", "income$25-$34,999", "income$35-$49,999", "income$50-$74,999", "income$75-$99,999", "income$100-$149,999", "income$150-$199,999", "income>$200", "median_household_income", "With_earnings", "mean_household_income")
colnames(davidson_income_census_2000)<-income_headers
```
```{r}

```
remove all commas from numerical data and mutate across to convert character values to numeric

```{r}
davidson_income_census_2000 <- davidson_income_census_2000 |> 
    mutate(across(-c(NAME, VALUE), ~as.numeric(str_remove_all(., ","))))
```

remove "with earnings" column
```{r}
davidson_income_census_2000 <- davidson_income_census_2000 |> 
    select(-With_earnings)
```

rename columns for 2010 and 2020:
```{r}
income_headers <- c("GEO_ID", "NAME", "HOUSEHOLDS", "income<$10k", "income$10-$14,999","income$15-$24,999", "income$25-$34,999", "income$35-$49,999", "income$50-$74,999", "income$75-$99,999", "income$100-$149,999", "income$150-$199,999", "income>$200", "median_household_income", "mean_household_income")
colnames(davidson_income_census_2010)<-income_headers

colnames(davidson_income_census_2020)<-income_headers
```
remove first row
```{r}
davidson_income_census_2010 <- davidson_income_census_2010 |>
    slice(-1)
```


```{r}

davidson_income_census_2020 <- davidson_income_census_2020 |> 
    slice(-1)

```
change all columns to numeric
```{r}
davidson_income_census_2010 <- davidson_income_census_2010 |> 
    mutate(across(-c(GEO_ID, NAME), ~as.numeric(.)))

davidson_income_census_2020 <- davidson_income_census_2020 |> 
    mutate(across(-c(GEO_ID, NAME), ~as.numeric(.)))
```
create counts from percent values for 2010 and 2020

```{r}
davidson_income_counts_2010 <- davidson_income_census_2010 |> 
    mutate(across(-c(GEO_ID, NAME, HOUSEHOLDS, median_household_income, mean_household_income),~round(HOUSEHOLDS*./100)))

davidson_income_counts_2020 <- davidson_income_census_2020 |> 
    mutate(across(-c(GEO_ID, NAME, HOUSEHOLDS, median_household_income, mean_household_income),~round(HOUSEHOLDS*./100)))
```

spot check rows to make sure calculations add up to tract total households
```{r}
davidson_income_counts_2010 |> 
    rowwise() |> 
    select(-c(GEO_ID, NAME, HOUSEHOLDS, median_household_income, mean_household_income)) |> 
    rowSums()
```
and double check:
```{r}
davidson_income_counts_2010 |> 
    pull(HOUSEHOLDS)
```



